\section{Introduction}

The past 6 years have seen the development of methodologies to identify genomic variation within a fetus through the sequencing of maternal blood plasma. These novel non-invasive methods are based on the observation that maternal plasma contains a fraction of DNA (typically 5-15\%) originating from the fetus \citep{lo1997presence}. The key advantage of non-invasive methods is that they pose no associated risk to the pregnancy. In contrast, invasive methods like chorionic villus sampling or amniocentesis (sampling of amniotic fluid from around the developing fetus) have estimated procedure-related fetal loss rate of 0.6\% to 1\%, \cite{douglas2007amnio}. Non-invasive methods have already been used, mainly for the detection of whole-chromosome events like trisomy of chromosome 21 (Down syndrome) \citep{chiu2008noninvasive, fan2008noninvasive}, and to a more limited extent for smaller (typically several megabases long) Copy Number Variants (CNVs) \citep{chen2013, srinivasan2013, rampasek2014fcnv}. Methods utilizing genome-wide detection of sub-chromosomal CNVs are highly desired, to enable prenatal screening for diseases like DiGeorge syndrome (\ntilde3Mb deletion), Prader-Willi syndrome (\ntilde4Mb deletion), and other, associated with a mid to large sized CNV. In addition to CNV detection, a successful proof-of-concept for non-invasive genome-wide fetal genotyping has been published \citep{kitzman2012}. This is the only work (to our best knowledge), in the field of maternal plasma analysis for fetal variation detection, with available data set.

In this project we will work on a new method for non-invasive fetal CNV detection given paternal genomes and deep sequencing of maternal plasma cell free DNA (cfDNA). We will build on \citep{rampasek2014fcnv}, seeking to addresses several of its current modelling drawbacks. Foremost, we will employ a discriminative model like a Conditional Random Field to focus on the classification goal and to better incorporate multiple dependant data features. We plan to investigate multiple ways of such CRF application, analysing the advantages and limitations of such models compared to Hidden Markov Model used in \citep{rampasek2014fcnv}. We will use the \textit{in silico} simulated CNV data set by \citep{rampasek2014fcnv} based on sequencing samples of \citep{kitzman2012}.

In the following sections, we give a brief description of the methods published so far, provide motivation for employing CRF rather than HMM for the CNV detection task, and outline the project goals.

\section{Existing methods}
To our best knowledge there are currently three methods for whole-genome detection of sub-chromosomal copy number variants in fetal genome from maternal plasma cfDNA sequencing.

\subsection{Methods utilizing depth-of-coverage signal}
Two methods published last year \citep{chen2013, srinivasan2013}, that try to address the aforementioned problem, utilize only low-coverage high-throughput sequencing of cfDNA from maternal plasma. Such sequencing is cheap (in order of few hundred USD per patient) what makes them well suited for practice in this regard.

While the exact methods used in both of these approaches differ, the essential idea remains the same. They map the reads to the reference genome, divide the genome into bins, and identify the CNVs by comparing the number of reads mapped to each bin with values obtained in control samples. The key idea in these methods is that deletions/duplications will result in more/fewer fetal reads within a window, and this difference can be identified using statistical methods.

In \cite{srinivasan2013} authors use depth-of-coverage computed in 1Mb windows across the genome to identify CNVs that are typically at least 1Mb long, though they do report discovery of a 300kb CNV. 9 of the 22 discovered CNVs in 11 patients were concordant with karyotyping results, with most discrepancies being short ($<$1Mb) CNVs.

In \cite{chen2013} authors use the same WG sequencing samples as are currently used for trisomy detection. Average DOC of such samples is less than one. Thus they use even larger 10Mb windows, again considering only the number of fragments mapped and are able to successfully identify variants 9-29Mb with only one false positive among 6 true positives in 1311 patients.

The limiting factor of these methods is the relatively large minimal CNV length  necessary for successful detection. However \cite{srinivasan2013} claims resolution down to 1Mb or even less, the paper provide only very limited justification, as only 3 from 16 calls of size $<$2Mb were true positives.

\subsection{Methods incorporating multiple signals}
In recent work \citep{rampasek2014fcnv} (currently in publication process) we introduced a probabilistic method for whole-genome non-invasive analysis of \textit{de novo} CNVs in fetal genome based on maternal plasma sequencing with largely increased sensitivity compared to methods published before. The method combines three types of information within a unified Hidden Markov Model: the imbalance of allelic ratios at SNP positions; the use of parental genotypes to phase nearby SNPs; and depth of coverage to better differentiate between various types of CNVs and to improve precision. Simulation results, based on \emph{in silico} introduction of novel CNVs into plasma samples with 13\% fetal DNA concentration, demonstrate a sensitivity of 90\% for CNVs $>$400 kilobases (with 13 calls in an unaffected genome), and 40\% for 50-400kb CNVs (with 108 calls in an unaffected genome).